The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-08-02 paper/paper.Rmd

\begin{equation}
\label{eq:model}
y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
\end{equation}

where $\beta_{0,j}$ is the diameter-independent growth rate of species $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census and $\beta_{\text{dbh},j}$ the slope of that species's diameter-growth relationship;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$. @allen_permutation_2020 estimate all parameters via Bayesian linear regression, while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004]. These closed-form solutions are not as computationally expensive as approximations from Markov Chain Monte Carlo algorithms.

@@

We load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2020/11/20 and perform minor data wrangling [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020]. We then only consider a 9 ha subsection of the 25.6 ha of the site to speed up computation for this example: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.

@@

@@

```{r appendix2-code}

census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = mdy(date),
    dbh = as.numeric(dbh)/10
  ) %>%
  filter(gx < 400, gy < 400)

@@

for (i in 1:length(mult_dist_comp$dist))
{
  comp_dist <- mult_dist_comp$dist[i]

@@

mult_dist_comp %>%
  ggplot(aes(dist,rmse)) +
  geom_point() +
  xlab('Distance (m)') +
  ylab('RMSE (cm)')

```

>>>>

\begin{equation}
\label{eq:model}
y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij}
\end{equation}

where $\beta_{0,j}$ is the diameter-independent growth rate of species $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census and $\beta_{\text{dbh},j}$ the slope of that species's diameter-growth relationship;  $x^{\text{comp}}_{ijk}$ is the sum of some numerical explanatory variable of all trees of competitor species $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.

@allen_permutation_2020 use the sum of the basal area of all trees of competitor species $k$ as $x^{\text{comp}}_{ijk}$. Furthermore, they estimate all parameters via Bayesian linear regression, while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004]. These closed-form solutions are not as computationally expensive as approximations from Markov Chain Monte Carlo algorithms.

@@

We load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2021/08/02 and perform minor data wrangling [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020]. We then only consider a 9 ha subsection of the 25.6 ha of the site to speed up computation for this example: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.

@@

We also load species information as it existed on GitHub on 2021/08/02, which includes family, genus, and species information; as well as classifications of the canopy position (canopy, canopy emergent, understory, shrub layer) and drought tolerance (intolerant, resistant) of the species.

```{r scbi-load-sp-data}
sp_info <- read_csv("SCBI_ForestGEO_sp_ecology.csv") %>%
  select(
    sp = spcode, family, genus, species, canopy_position, 
    drought_tolerance
  ) %>% 
  mutate()
sp_info
```

We join this species information to our `growth_scbi` data frame.

```{r}
growth_scbi <- growth_scbi %>% 
  left_join(sp_info, by = "sp")
```

Furthermore, we compute two potential competitor explanatory variables $x^{\text{comp}}_{ijk}$ from Equation \ref{eq:model}. First, the basal area of each tree as a function of its DBH in the earlier census. Second, the above ground biomass as estimated by allometric equations encoded in the `get_biomass()` function from the `allodb` package [@allodb_package]; this function has DBH, species, and geographic coordinates as arguments.

```{r scbi-compute-comp-explanatory-variables}
# Install development version of allodb using:
# remotes::install_github("forestgeo/allodb")
library(allodb)
growth_scbi <- growth_scbi %>% 
  mutate(
    # Compute basal area:
    basal_area = 0.0001 * pi * (dbh1 / 2)^2,
    # Compute above ground biomass:
    agb = get_biomass(
      dbh = dbh1,
      genus = genus,
      species = species,
      coords = c(-78.2, 38.9)
    )
  )
```

@@

```{r appendix2-code, eval = FALSE}
census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = mdy(date),
    dbh = as.numeric(dbh)/10
  ) %>%
  filter(gx < 400, gy < 400)

@@

for (i in 1:length(mult_dist_comp$dist)){
  comp_dist <- mult_dist_comp$dist[i]

@@

mult_dist_comp %>%
  ggplot(aes(dist,rmse)) +
  geom_point() +
  xlab('Distance (m)') +
  ylab('RMSE (cm)')
```
